---
title: "Test - Assignment 2"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)
library(summarytools)
library(kableExtra)
library(performance)

##### initialise datasets
dat <- read.csv('Ass21Data.csv', header = TRUE, na.strings = c('NA',' N/A'), stringsAsFactors = TRUE)

##### global parameters
SEED = 2026
```



```{r cv_with_glmnet, fig.width=10, fig.height=5}

library(glmnet)
SEED = 2026
lambda_grid <- 10^seq(3, -4, length.out = 100)
##### initialise
# summary
# summary(dat)
# define trainset and testset
train_dat <- dat[dat$OBS_TYPE == "Train", ]
test_dat  <- dat[dat$OBS_TYPE == "Test", ]
# model formula
formula_death <- DEATH_RATE ~ . - CODE - OBS_TYPE
# matrix prep (required for glmnet)
x_train <- model.matrix(formula_death, data = train_dat)[,-1]
y_train <- train_dat$DEATH_RATE


##### CV for Ridge/Lasso
# run CV
set.seed(SEED)
best_cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, family = "gaussian")
set.seed(SEED)
best_cv_lasso <- cv.glmnet(x_train, y_train, alpha = 1, family = "gaussian")
# extract best lambdas
min_lambda_ridge <- best_cv_ridge$lambda.min
cat("Best Ridge Lambda (min):", min_lambda_ridge, "\n")
ose_lambda_ridge <- best_cv_ridge$lambda.1se
cat("1se Ridge Lambda (1se):", ose_lambda_ridge, "\n")
min_lambda_lasso <- best_cv_lasso$lambda.min
cat("Best Lasso Lambda (min):", min_lambda_lasso, "\n")
ose_lambda_lasso <- best_cv_lasso$lambda.1se
cat("1se Lasso Lambda (1se):",ose_lambda_lasso, "\n")
ose_lambda_lasso
# CV error curve
par(mfrow = c(1, 2), mar = c(5, 4, 8, 2))
plot(best_cv_ridge, main = "Ridge: CV Error vs Lambda")
plot(best_cv_lasso, main = "LASSO: CV Error vs Lambda")


##### CV for elastic net
alpha_grid <- seq(0, 1, by = 0.1)
# CV function (elastic net)
cv_elastic <- lapply(alpha_grid, function(a) {
  set.seed(SEED)
  cv <- cv.glmnet(x_train, y_train,
                  alpha  = a,
                  lambda = lambda_grid,
                  nfolds = 10)
  list(alpha      = a,
       cv         = cv,
       lambda_min = cv$lambda.min,
       lambda_1se = cv$lambda.1se,
       cvm_min    = min(cv$cvm))
  })
# CV index with best alpha
best_cv_idx <- which.min(sapply(cv_elastic, function(x) x$cvm_min))
best_alpha  <- cv_elastic[[best_cv_idx]]$alpha
best_cv_elastic <- cv_elastic[[best_cv_idx]]$cv
cat("Best Elastic Net Alpha: ", best_alpha, "\n")
# extract best lambdas
min_lambda_elastic <- best_cv_elastic$lambda.min
cat("Best Elastic Net Lambda (min):", min_lambda_elastic, "\n")
ose_lambda_elastic <- best_cv_elastic$lambda.1se
cat("1se Elastic Net Lambda (1se):", ose_lambda_elastic, "\n")
# CV error curve for best alpha
par(mfrow = c(1, 1), mar = c(5, 4, 8, 2))
plot(best_cv_elastic, main = paste0("ElasticNet CV: Error vs Lambda (alpha = ", best_alpha, ")"))

```




```{r test_3, fig.width=3.5, fig.height=3.5}
##### summary
summary(dat)


##### define trainset and testset
train_dat <- dat[dat$OBS_TYPE == "Train", ]
test_dat  <- dat[dat$OBS_TYPE == "Test", ]


##### define the model formulas
# model formula
formula_death <- DEATH_RATE ~ . - CODE - OBS_TYPE
# all predictor names in formula
colnames(model.matrix(formula_death, data = train_dat))


##### summary
# define specific hyper-parameters
grid_of_interest <- expand.grid(alpha = 0, lambda = 0.129155)
# run the model
mod_gaussian_caret <- caret::train(
  formula_death, 
  data = train_dat, 
  method = "glmnet",
  family = "gaussian",
  tuneGrid = grid_of_interest,
  trControl = trainControl(method = "none") # "none" means not to cv
)


##### view the final coefficients
coef(mod_gaussian_caret$finalModel, mod_gaussian_caret$bestTune$lambda)


##### diagnostic plots for the GLM
pred_test <- predict(mod_gaussian_caret, newdata = test_dat)
res_test <- test_dat$DEATH_RATE - pred_test

plot(pred_test, res_test,
     xlab = "Fitted values",
     ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red")

qqnorm(res_test)
qqline(res_test, col = "red")

plot(pred_test, sqrt(abs(res_test)),
     xlab = "Fitted values",
     ylab = "Sqrt(|Residuals|)",
     main = "Scale-Location")


```




```{r cv_bug_code, fig.width=10, fig.height=5}
library(recipes)

df <- dat  # full data as app uses transform$data

rec <- recipe(DEATH_RATE ~ ., data = df) |>
  update_role(CODE,     new_role = "ignore") |>
  update_role(OBS_TYPE, new_role = "ignore")

prepped <- prep(rec, training = df, verbose = FALSE)
baked   <- bake(prepped, new_data = df)

cat("Baked columns:\n")
cat(names(baked), sep = "\n")
cat("\nN cols:", ncol(baked), "\n")

# what get_xy does
x_df <- baked[, setdiff(names(baked), "DEATH_RATE"), drop = FALSE]
cat("\nX cols after setdiff:", ncol(x_df), "\n")
cat(names(x_df), sep = "\n")
```


<div class="question">

## **Question 1**

</div>

\vspace{2em}




























<div class="question">

## **Reproducibility Statement**

</div>

All analyses were performed in R using the packages readxl, tidyverse, summarytools, kableExtra, corrplot, car, seriation, MASS, and glmnet. The datasets used were Residen.RData, Residential-Building-Data-Set.xlsx, parkinsons.csv, and Weather_Station_Data_v1.csv as provided in the assignment materials. A global random seed of 2026 (SEED = 2026) was set throughout to ensure reproducibility of all train–test splits and cross-validation procedures. Full session information is recorded in the sessionInfo() output below.

```{r session_info}
# sessionInfo for reproducibility
sessionInfo()
```

<div class="question">

## **Appendix: AI Use Declaration**

</div>

GPT assisted in:

- Structuring this R Markdown document by defining global headers, global helper functions, enabling code chunks to appear more compact and improving the overall visual organization of the document;
- Providing LaTeX syntax for mathematical expressions so that equations could be properly represented;
- Explaining Test Error decomposition in terms of bias and variance;
- Providing syntax explaination for glmnet();
- Assisting with general debugging.
- Assisting with accurate academic wording to help express clearly.

<div class="question">

## **References**

</div>

<blank>