---
title: "DATA423 - Tutorial_on_Resampling_sample_code_rmd"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)


##### initialise datasets
# dat <- read.csv('Ass3Data.csv', header = TRUE, na.strings = c('NA',' N/A'), stringsAsFactors = TRUE)

##### global parameters
SEED = 2026
```

```{r global_helper_comparison_table}
##### blablabla
# blabla placeholder
```



```{r f0, fig.width=10, fig.height=10}
##### Introduction

library(caret)

# Independent Observartions
#   Simple Random Sampling
#   Simple Cross-Validation
#   Stratified Random sampling
#   Simple Bootstrap sampling
#   Stratified Bootstrap sampling

# Group Dependent Observations
#   Leave Group Out Sampling

# Time Dependent Observations
#   Time Series Sampling

# Down Sampling with diversity
#   Diversity Down-Sampling

```


```{r f1, fig.width=10, fig.height=10}
##### Data

data(iris, package = "datasets")
summary(iris)


```




```{r f2, fig.width=10, fig.height=10}
##### Independent Observartions - Sampling into 2 partitions (train + test)

# the frequencies of the outcome variable Species can be tallied (or tabulated) using the table() function
table(iris$Species)

# a Simple Random Sampling not guaranteed with original class proportion
set.seed(61)
N <- nrow(iris)
trainIndex <- base::sample(x = N, size = 0.8 * N, replace = FALSE)
table(iris[trainIndex, "Species"])

# a Stratified Random Sampling can do better by preserving class balance
#   for classification, stratification is straightforward
#   for regression, artificial groups using quartiles
set.seed(61)
trainIndex <- caret::createDataPartition(y = iris$Species, p = 0.8, list = FALSE)
table(iris[trainIndex, "Species"])

```


```{r f3, fig.width=10, fig.height=10}
##### Independent Observartions - Sampling into 3 partitions (train + val + test)
set.seed(61)
subIndex <- caret::createDataPartition(y = iris$Species, p = 0.6, list = FALSE)
train <- iris[subIndex,]
s40 <- iris[-subIndex,]
cat("There are", nrow(train), "observations in the train data")
subIndex <- caret::createDataPartition(y = s40$Species, p = 15/40, list = FALSE)
validation <- s40[-subIndex,]
cat("There are", nrow(validation), "observations in the validation data")
test <- s40[subIndex,]
cat("There are", nrow(test), "observations in the test data\n")

```



```{r f4, fig.width=10, fig.height=10}
##### Independent Observartions - Stratified Bootstrap Sampling

data(BostonHousing, package = "mlbench")

# createResample() function performs stratified bootstrapping based on the outcome variable
set.seed(61)
resamples <- caret::createResample(y = BostonHousing$medv, times = 15, list = TRUE)
str(resamples)

# compute the actual proportion of unique observations in the resamples
calcPerc <- function(x) {
  x %>% 
    unique() %>% 
    length() -> unqCount
  propUnq <- unqCount / length(x) 
  cat("Unique =", round(propUnq * 100), "%\n")
}
invisible(mapply(resamples, FUN = calcPerc))

```




```{r f5, fig.width=10, fig.height=10}
# Group Dependent Observations - Leave Group Out Sampling

# create some artificial data
set.seed(357)
patients <- sample(LETTERS, size = 100, replace = TRUE)
patients

table(as.factor(patients))

# Patient T has three observations while patient X has 9 observations
# a good splitting strategy would be to create 5 cross-validation folds using the patient identity as the “group”
set.seed(135)
folds <- caret::groupKFold(group = patients, k = 5) 
folds

# the parameter k sets the number of folds
# since we have 26 levels for patients, we can expect 265/5 letters to be held-out (on average)
inGroups <- sort(patients[folds$Fold1])
inGroups

# there are 26 possible letters, how many are present in fold1
distinct <- length(unique(inGroups))
cat("Distinct groups in train for fold1:", distinct)

# the hold-out set relating to fold1 would be the test data
outGroups <- sort(patients[-folds$Fold1])
outGroups

# how many letters (i.e. patients) are present in the hold-out set relating to fold1
distinct <- length(unique(outGroups))
cat("Distinct groups in test for fold1:", distinct)


```



```{r f6, fig.width=10, fig.height=10}
# Time Dependent Observations - Time Series Sampling

# need to carefully research the parameters initialWindow, horizon, and fixedWindow to achieve the sampling you want
tsSamples <- caret::createTimeSlices(y = 1:25, initialWindow = 5, horizon = 3, fixedWindow = FALSE)
print(str(tsSamples$train))
print(str(tsSamples$test))

```



```{r f7, fig.width=10, fig.height=10}
# Down Sampling with Diversity - Diversity Down-Sampling

# K-Mediods Method
features <- BostonHousing[, c("age", "nox")]
clusters <- cluster::pam(features, k = 80, metric = "euclidean", stand = TRUE)
trainIndex <- clusters$id.med  # the ids of the medoids of the clusters 
BostonHousing$Type <- NA
BostonHousing$Type[-trainIndex] <- "not sampled"
BostonHousing$Type[trainIndex] <- "sampled"
ggplot() +
  geom_point(data = BostonHousing, mapping = aes(x = age, y = nox, color = Type)) +
  labs(title = "Boston Data (sampled for diversity)")

# an exercise, generalised to use all of the BostonHousing variables
features <- BostonHousing
clusters <- cluster::pam(features, k = 80, metric = "euclidean", stand = TRUE)
trainIndex <- clusters$id.med  # the ids of the medoids of the clusters 
BostonHousing$Type <- NA
BostonHousing$Type[-trainIndex] <- "not sampled"
BostonHousing$Type[trainIndex] <- "sampled"
ggplot() +
  geom_point(data = BostonHousing, mapping = aes(x = age, y = nox, color = Type)) +
  labs(title = "Boston Data (sampled for diversity)")

```


# Wrap Up

The process of extracting train and test datasets is a sampling process.
The process of down-sampling data is also a sampling process.
The correct sampling strategy is not always Simple Random sampling.
Using the wrong sampling strategy can lead to Data Leakage.
The key to deciding on the best sampling strategy is to consider the dependencies (or lack of independence) between the observations.
