---
title: "DATA423 - Tutorial_on_Class_Imbalance_sample_code_rmd"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)


##### initialise datasets
# dat <- read.csv('Ass3Data.csv', header = TRUE, na.strings = c('NA',' N/A'), stringsAsFactors = TRUE)

##### global parameters
SEED = 2026
```

```{r global_helper_comparison_table}
##### blablabla
# blabla placeholder
```



```{r f0, fig.width=10, fig.height=10}
##### Introduction

library(caret)

# several approaches to dealing with class imbalance
#   Rebalancing (data-level fix, change the dataset itself)
#     the process of creating and deleting observations 
#     either upsampling (duplicate or synthesize) or downsampling (remove data)
#     risk: upsampling -> overfitting; downsampling -> lose information
#   Coalescing (label-level fix, change the labels)
#     the process of reassigning the target categories more evenly 
#     from A, B, C, D (multi-class) to {A, B} vs {C, D}
#     cardinality merging
#     risk: lose granularity
#   Cost-sensitive methods (objective-level fix, penalize mistakes differently)
#     models are not distracted by the more populous target levels 
#     make the model care more about minority errors
#   Weighted methods (data weighting, assign weights to observations)
#     methods that weight observations so as to ensure that models are not distracted by the more populous target levels
#     similar to rebalancing, but cleaner
#   Appropriate metrics (evaluation fix, change how you measure performance)
#     model metrics that are not distracted by the more populous target levels
#     do not use Accuracy
#   Anomaly detection (problem reformulation)
#     treat minority class as anomalies
#     turn the problem, fully or partly, into an anomaly detection problem
#     instead of classification, detect deviations from normal
```


```{r f1, fig.width=10, fig.height=10}
##### Method Selection - Null model

# regression null: attributes everything on a constant intercept, without any feature
# classification null: a most frequent class


```




```{r f2, fig.width=10, fig.height=10}
##### Method Selection - different methods compared to NULL

# define mutual split, and resampling strategy
trIndex <- caret::createDataPartition(y = iris$Species, p = 0.7, list = FALSE)
train <- iris[trIndex,]
control <- trainControl(method = "repeatedcv", number = 10, repeats = 2, classProbs = TRUE, verboseIter = TRUE)

# Null Method
NullModel <- caret::train(Species ~ ., data = train, method = "null", trControl = control)
NullModel

# Rpart2 Method
Rpart2Model <- caret::train(Species ~ ., data = train, method = "rpart2", trControl = control)
Rpart2Model

# Random Forest Method
RfModel <- caret::train(Species ~ ., data = train, method = "rf", trControl = control)
RfModel


```


```{r f3, fig.width=10, fig.height=10}
##### Method Selection - Comparison Charts 1

# visual difference
list("null" = NullModel, "rpart2" = Rpart2Model, "rf" = RfModel) %>%
resamples() %>%
bwplot(notch = FALSE)

# statistically difference
list("null" = NullModel, "rpart2" = Rpart2Model, "rf" = RfModel) %>%
resamples() %>%
bwplot(notch = TRUE)

```



```{r f4, fig.width=10, fig.height=10}
##### Method Selection - Comparison Charts 2

# multiclass metrics difference
control <- trainControl(method = "repeatedcv", number = 10, repeats = 2, classProbs = TRUE, summaryFunction = multiClassSummary)
myMetric <- "AUC"
NullModel <- caret::train(Species ~ ., data = train, method = "null", trControl = control, metric = myMetric)
Rpart2Model <- caret::train(Species ~ ., data = train, method = "rpart2", trControl = control, metric = myMetric)
RfModel <- caret::train(Species ~ ., data = train, method = "rf", trControl = control, metric = myMetric)
models <- list("null" = NullModel, "rpart2" = Rpart2Model, "rf" = RfModel)
bwplot(resamples(models), notch = TRUE)

```



# Wrap Up

Whatever choices you settle on for the train control specification, it is important that this is used consistently for all the candidate models being trained. In other words if you want to compare models fairly they should have the same Train-Control specification.
The null model provides a sense of scale by which we can judge visually whether we doing well or not.
Each model’s metric distribution can be reviewed. Outliers are a worry. The narrower the boxplot, the less uncertainty there is.
Notches are a means of informally checking whether the boxplot means/medians are statistically different. The notch covers the 95 percentile range of the mean.
Any models whose metric is worse than the null model is poorer than the simplest possible model. That is, it is truly useless.
The available metrics are determined by the selectionFunction value.
The metric that drives the hyper-parameter optimisation can be supplied in the train() function’s metric parameter - see last example.
