---
title: "DATA423 - caret Learning"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)


##### initialise datasets
# load("Residen.RData")
# parkinsons <- read.csv("parkinsons.csv")

##### global parameters
SEED = 2026
```

```{r global_helper_comparison_table}
##### function to plot correlation heatmap based on dataset
# blabla placeholder
```

<div class="question">

## **Function 1**

featureplot()

</div>

\vspace{2em}

### F1 Code

```{r f1, fig.width=10, fig.height=10}
##### blabla

library(caret)
data(iris)
str(iris)
featurePlot(x=iris[,1:4], y=iris$Species, plot='pairs')


```




<div class="question">

## **Function 2**

preprocess(): imputation

Note: RANN is a dependency needed for knnImpute in caret

</div>

\vspace{2em}

### F2 Code

```{r f2, fig.width=10, fig.height=10}
##### blabla

Titanic <- read.csv("https://raw.githubusercontent.com/mariocastro73/ML2020-2021/master/datasets/Titanic.csv")
age.mean <- mean(Titanic$Age, na.rm = TRUE)
age.sd <- sd(Titanic$Age, na.rm = TRUE)
pre <- preProcess(Titanic, method = "knnImpute", k = 5)
Titanic.imputed <- predict(pre, Titanic)
Titanic$Age[10:15]
Titanic.imputed$Age[10:15]*age.sd+age.mean


```



<div class="question">

## **Function 3**

preprocess(): dummy

Dummy Vars

When dummy encoding with k levels, we have two choices:
- Without full rank
- - create linear dependence (Brown + Blonde + Black + Red + Grey = 1)
- - Perfect multicollinearity (dummy variable trap)
- With full rank
- - create k - 1 dummy variables (fullRank = TRUE)
- - Matrix is invertible (important for OLS)

</div>

\vspace{2em}

### F3 Code

```{r f3, fig.width=10, fig.height=10}
##### blabla

data <- data.frame(vars=as.factor(c("Brown","Blonde","Black","Red","Grey")))
dv <- dummyVars(~vars, data=data, fullRank = T)
cbind(data, predict(dv, data))


```




<div class="question">

## **Function 4**

preprocess(): nearZeroVar

nearZeroVar

</div>

\vspace{2em}

### F4 Code

```{r f4, fig.width=10, fig.height=10}
##### blabla

data(mdrr)
dim(mdrrDescr)
nzv <- nearZeroVar(mdrrDescr, saveMetrics= TRUE)
filteredDescr <- mdrrDescr[, !nzv$nzv]
dim(filteredDescr)


```



<div class="question">

## **Function 5**

createDataPartition()

splitting

</div>

\vspace{2em}

### F5 Code

```{r f5, fig.width=10, fig.height=10}
##### blabla

data(Default, package = "ISLR")
dim(Default)
default_idx = createDataPartition(Default$default, p = 0.75, list = FALSE)
default_trn = Default[default_idx, ]
default_tst = Default[-default_idx, ]
str(default_trn) # Training

str(default_tst) # Testing

```



<div class="question">

## **Function 6**

train()
predict()

training

</div>

\vspace{2em}

### F6 Code

```{r f6, fig.width=10, fig.height=10}
##### blabla

# Load the dataset directly from URL
admision <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

# This dataset has exactly: admit, gre, gpa, rank
str(admision)

# Split
set.seed(42)
idx <- createDataPartition(admision$admit, p = 0.75, list = FALSE)
admision.trn <- admision[idx, ]
admision.tst <- admision[-idx, ]

# Then your original chunk works unchanged:
admision$admit <- as.factor(admision$admit)
admision.trn$admit <- as.factor(admision.trn$admit)
admision.tst$admit <- as.factor(admision.tst$admit)

fit.logistic <- train(admit ~ ., data = admision.trn, method = "glm", family = "binomial")
fit.knn      <- train(admit ~ ., data = admision.trn, method = "knn")
pred.logistic <- predict(fit.logistic, admision.tst)
pred.knn      <- predict(fit.knn, admision.tst)
table(admision.tst$admit, pred.logistic) # More accurate than knn
table(admision.tst$admit, pred.knn)

```



<div class="question">

## **Function 6**

resamples()

model selecting

Kappa is a measure of the difference between observed and expected

- standardized to lie on a -1 to 1 scale, where 1 is perfect agreement
- 0 is exactly what would be expected by chance
- negative values indicate agreement less than chance

</div>

\vspace{2em}

### F6 Code

```{r f6, fig.width=8, fig.height=5}
##### blabla

library(mlbench) # Library to compare fittings with different methods
results <- resamples(list(LR=fit.logistic, KNN=fit.knn))
dotplot(results)


```






## **Appendix: AI Use Declaration**

</div>

GPT bla

<div class="question">

## **References**

</div>

<blank>