---
title: "DATA423 - caret Learning"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)
library(caret)

##### initialise datasets
dat <- read.csv('Ass3Data.csv', header = TRUE, na.strings = c('NA',' N/A'), stringsAsFactors = TRUE)

##### global parameters
SEED = 2026
```

```{r global_helper_comparison_table}
##### function to plot correlation heatmap based on dataset
# blabla placeholder
```

<div class="question">

## **Splitting Based on Categorical Outcome**

featureplot()

</div>

\vspace{2em}

### F1 Code

```{r f1, fig.width=10, fig.height=10}
##### assuming blood type is that categorical outcome
# list = FALSE, avoids returning the data as a list
# times, can create multiple splits at once


set.seed(SEED)
train_idx_cat <- createDataPartition(dat$BloodType, p = .8, 
                                  list = FALSE, 
                                  times = 1)

head(train_idx_cat)

# train_cat_y <- dat[train_idx_cat, ]
# test_cat_y  <- dat[-train_idx_cat, ]

```



<div class="question">

## **Splitting Based on the Predictors**

maxDissim()

</div>

\vspace{2em}

### F2 Code

```{r f2, fig.width=10, fig.height=10}
##### 
library(mlbench)
data(BostonHousing)

testing <- scale(BostonHousing[, c("age", "nox")])
set.seed(5)
## A random sample of 5 data points
startSet <- sample(1:dim(testing)[1], 5)
samplePool <- testing[-startSet,]
start <- testing[startSet,]
newSamp <- maxDissim(start, samplePool, n = 20)
head(newSamp)

```



<div class="question">

## **Data Splitting for Time Series**

createTimeSlices

</div>

\vspace{2em}

### F3 Code

```{r f3, fig.width=10, fig.height=10}
##### 

# treat data as date
dat$ObservationDate <- as.Date(dat$ObservationDate)

# Sort by time first - critical for time series splitting
dat <- dat[order(dat$ObservationDate), ]

# Create time slices
slices <- createTimeSlices(
  y             = seq_len(nrow(dat)),  # just indices
  initialWindow = 200,   # train on first 200 obs
  horizon       = 50,    # predict next 50
  fixedWindow   = FALSE  # growing window
)

# Example: train/test on the first fold
train_idx <- slices$train[[1]]
test_idx  <- slices$test[[1]]

train_dat <- dat[train_idx, ]
test_dat  <- dat[test_idx, ]

# Quick check
cat("Training rows:", nrow(train_dat), "\n")
cat("Test rows:",     nrow(test_dat),  "\n")
cat("Train period:", as.character(range(train_dat$ObservationDate)), "\n")
cat("Test period:",  as.character(range(test_dat$ObservationDate)),  "\n")

# Fit a model on one fold (example with lm)
model <- train(
  Response ~ Alcohol + Coffee + Exercise + ChemoTreatments + BloodType,
  data      = train_dat,
  method    = "lm",
  na.action = na.omit
)

pred <- predict(model, newdata = test_dat)
RMSE <- sqrt(mean((pred - test_dat$Response)^2, na.rm = TRUE))
cat("Test RMSE:", RMSE, "\n")

```



<div class="question">

## **Simple Splitting with Important Groups**

groupKFold

</div>

\vspace{2em}

### F4 Code

```{r f4, fig.width=10, fig.height=10}
##### 

set.seed(SEED)
subjects <- sample(1:20, size = 80, replace = TRUE)
table(subjects)

# the results in folds can be used as inputs into the index argument of the trainControl function
folds <- groupKFold(subjects, k = 15)

```
