---
title: "DATA423 - Tutorial_on_Method_Selection_sample_code_rmd"
author: |
  William Hui Chang (69051925)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
header-includes:
- \usepackage{xcolor}
- \usepackage{etoolbox}
- \AtBeginEnvironment{verbatim}{\scriptsize}
- \AtBeginEnvironment{Shaded}{\scriptsize}
---

<style>
body { background-color: #ffffff; }
.question { 
  background-color: #2C3E50;
  color: white; 
  padding: 12px 15px; 
  border-radius: 4px; 
  margin: 15px 0;
  font-size: 1.4em;
}
</style>

```{r setup, include=TRUE, message = FALSE, warning=FALSE}
##### set global default for all plots
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.width=5.5, fig.height=3.5, fig.align='left')

##### load libraries
library(tidyverse)


##### initialise datasets
# dat <- read.csv('Ass3Data.csv', header = TRUE, na.strings = c('NA',' N/A'), stringsAsFactors = TRUE)

##### global parameters
SEED = 2026
```

```{r global_helper_comparison_table}
##### function to plot correlation heatmap based on dataset
# blabla placeholder
```



```{r f0, fig.width=10, fig.height=10}
##### Introduction

library(caret)

```


```{r f1, fig.width=10, fig.height=10}
##### Method discovery - a complex list of lists

mi <- caret::getModelInfo()
print(names(mi))

```



```{r f2, fig.width=10, fig.height=10}
##### Method discovery - What properties does each method record?

names(mi[[1]]) #note double [[]]

# for now we are interested in the short name, long name, type, description-tags and packages
names(mi[1]) # note single []
mi[[1]]$label
mi[[1]]$type
mi[[1]]$tags
mi[[1]]$library


```



```{r f3, fig.width=10, fig.height=10}
##### Method discovery - transform the method catalogue into a data frame
caretMethodsDF <- function(useHTML = FALSE) {
  mi <- caret::getModelInfo()
  if (useHTML) {
    breaker <- "<br>"
  } else {
    breaker <- "\n"
  }
  
  Label <- vector(mode = "character", length = length(mi))
  Packages <- vector(mode = "character", length = length(mi))
  Hyperparams <- vector(mode = "character", length = length(mi))
  Regression <- vector(mode = "logical", length = length(mi))
  Classification <- vector(mode = "logical", length = length(mi))
  Tags <- vector(mode = "character", length = length(mi))
  ClassProbs <- vector(mode = "logical", length = length(mi))
  for (row in 1:length(mi)) {
    Label[row] <- mi[[row]]$label
    libs <- mi[[row]]$library
    libs <- na.omit(libs[libs != ""]) # remove blank packages
    if (length(libs) > 0) {
      Packages[row] <- paste(collapse = breaker, mi[[row]]$library)
    }
    d <- mi[[row]]$parameters
    Hyperparams[row] <- paste(collapse = breaker, paste0(d$parameter, " - ", d$label, " [", d$class,"]"))
    Regression[row] <- "Regression" %in% mi[[row]]$type
    Classification[row] <- "Classification" %in% mi[[row]]$type
    Tags[row] <- paste(collapse = breaker, mi[[row]]$tags)
    ClassProbs[row] <- is.function(mi[[row]]$prob)
  }
  data.frame(Model = names(mi), Label, Packages, Regression, Classification, Tags, Hyperparams, ClassProbs, stringsAsFactors = FALSE)
}  


df <- caretMethodsDF()
df


```




```{r f4, fig.width=10, fig.height=10}
##### Method discovery - a better df, html style
df <- caretMethodsDF(useHTML = TRUE)
DT::datatable(df, escape = FALSE, rownames = FALSE, extensions = list(Responsive = TRUE))


```




```{r f5, fig.width=10, fig.height=10}
##### Visualising Methods - utility in turning the Tags into a wide data frame

# make the relevant data numeric and 
# we can calculate distances between methods (i.e. similarity / dissimilarity) based on each tag
modelInfo <- caret::getModelInfo()
tags <- vector(mode = "list", length = length(modelInfo))
Classification <- Regression <- ClassProbs <- rep(NA, length = length(modelInfo))
for (i in seq(along = modelInfo)){
  tags[[i]] <- modelInfo[[i]]$tags
  Classification[i] <- ifelse("Classification" %in% modelInfo[[i]]$type, 1, 0)
  Regression[i] <- ifelse("Regression" %in% modelInfo[[i]]$type, 1, 0)
  ClassProbs[i] <- ifelse(is.null(modelInfo[[i]]$prob), 0, 1)
}
tabs <- table(unlist(tags))
tabs <- tabs[order(tolower(names(tabs)))]
terms <- names(tabs)
terms <- terms[terms != ""]
dat <- matrix(0, ncol = length(terms), nrow = length(tags))
colnames(dat) <- terms
hasTag <- lapply(tags, function(x, y) which(y %in% x), y = terms)
for (i in seq(along = hasTag)) {
  dat[i, hasTag[[i]]] <- 1
}
dat <- cbind(Classification, Regression, dat)
wide <- as.data.frame(dat, row.names = names(modelInfo))
wide


```



```{r f6, fig.width=10, fig.height=10}
##### Visualising Methods - Classification Map

# we can restrict ourselves to classification methods only
wide[wide$Classification == 1, ] %>%
  stats::dist(method = "manhattan") -> d   # other choices are "euclidean", "binary", "canberra"
print("Classification distance-matrix produced")

# a dimensional reduction to 2 features so we can plot this as a 2D plot
# cmdscale() function is just one of many ways to dimensionally reduce this data
d %>%
  stats::cmdscale(k = 2) %>%
  data.frame() -> dd

# plot this 2 dimensional data, a map of classification methods
ggplot(data = dd, mapping = aes(x = X1, y = X2, label = rownames(dd))) +
  ggtitle(label = "Classification Methods") +
  xlab("Coordinate 1") +
  ylab("Coordinate 2") +
  geom_point() +
  ggrepel::geom_text_repel(size = 2, max.overlaps = 50, na.rm = TRUE) +
  theme(plot.title = element_text(lineheight = 1, face = "bold", hjust = 0.5))


# the same for regression methods, this time trying euclidean distance with the stats::dist() function
# a map of regression methods
wide[wide$Regression == 1,] %>%
  stats::dist(method = "euclidean") %>%    # other choices are "euclidean", "binary", "canberra"
  stats::cmdscale(k = 2) %>%
  data.frame() -> dd

ggplot(data = dd, mapping = aes(x = X1, y = X2, label = rownames(dd))) +
  ggtitle(label = "Regression Methods") +
  xlab("Coordinate 1") +
  ylab("Coordinate 2") +
  geom_point() +
  ggrepel::geom_text_repel(size = 2, max.overlaps = 50, na.rm = TRUE) +
  theme(plot.title = element_text(lineheight = 1, face = "bold", hjust = 0.5))

```



```{r f7, fig.width=10, fig.height=10}
##### Filtering methods - 
# suppose we want to 
#   select for regression methods, and 
#   against linear methods, and 
#   against robustness, but 
#   for ensembling

df <- caretMethodsDF()
matches <- df$Regression == TRUE & !grepl(df$Tags, pattern = "Linear", ignore.case = TRUE) & !grepl(df$Tags, pattern = "\\bRobust Model\\b") & grepl(df$Tags, pattern = "\\bEnsemble Model\\b")
df[matches,]

```



List of tags

Accepts Case Weights
Bagging
Bayesian Model
Binary Predictors Only
Boosting
Categorical Predictors Only
Cost Sensitive Learning
Discriminant Analysis
Distance Weighted Discrimination
Ensemble Model
Feature Extraction
Feature Selection Wrapper
Gaussian Process
Generalized Additive Model
Generalized Linear Model
Handle Missing Predictor Data
Implicit Feature Selection
Kernel Method
L1 Regularization
L2 Regularization
Linear Classifier
Linear Regression
Logic Regression
Logistic Regression
Mixture Model
Model Tree
Multivariate Adaptive Regression Splines
Neural Network
Oblique Tree
Ordinal Outcomes
Partial Least Squares
Patient Rule Induction Method
Polynomial Model
Prototype Models
Quantile Regression
Radial Basis Function
Random Forest
Regularization
Relevance Vector Machines
Ridge Regression
Robust Methods
Robust Model
ROC Curves
Rule-Based Model
Self-Organising Maps
String Kernel
Support Vector Machines
Supports Class Probabilities
Text Mining
Tree-Based Model
Two Class Only